---
title: "Homework 6"
author: "Shelley Shen"
date: "12/7/2020"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(modelr)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

## Problem 1

```{r}
homicide_df = 
  read_csv("data/homicide-data.csv", na = c("", "NA", "Unknown")) %>% 
  mutate(
    city_state = str_c(city, state, sep = ", "),
    victim_age = as.numeric(victim_age),
    resolution = case_when(
      disposition == "Closed without arrest" ~ 0,
      disposition == "Open/No arrest"        ~ 0,
      disposition == "Closed by arrest"      ~ 1)
  ) %>% 
  filter(
    victim_race %in% c("White", "Black"),
    city_state != "Tulsa, AL") %>% 
  select(city_state, resolution, victim_age, victim_race, victim_sex)
```

Start with one city.

```{r}
baltimore_df =
  homicide_df %>% 
  filter(city_state == "Baltimore, MD")

glm(resolution ~ victim_age + victim_race + victim_sex, 
    data = baltimore_df,
    family = binomial()) %>% 
  broom::tidy() %>% 
  mutate(
    OR = exp(estimate),
    CI_lower = exp(estimate - 1.96 * std.error),
    CI_upper = exp(estimate + 1.96 * std.error)
  ) %>% 
  select(term, OR, starts_with("CI")) %>% 
  knitr::kable(digits = 3)
```

Try this across cities.

```{r}
models_results_df = 
  homicide_df %>% 
  nest(data = -city_state) %>% 
  mutate(
    models = 
      map(.x = data, ~glm(resolution ~ victim_age + victim_race + victim_sex, data = .x, family = binomial())),
    results = map(models, broom::tidy)
  ) %>% 
  select(city_state, results) %>% 
  unnest(results) %>% 
  mutate(
    OR = exp(estimate),
    CI_lower = exp(estimate - 1.96 * std.error),
    CI_upper = exp(estimate + 1.96 * std.error)
  ) %>% 
  select(city_state, term, OR, starts_with("CI")) 
```

``` {r}
models_results_df %>% 
  filter(term == "victim_sexMale") %>% 
  mutate(city_state = fct_reorder(city_state, OR)) %>% 
  ggplot(aes(x = city_state, y = OR)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = CI_lower, ymax = CI_upper)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```


## Problem 2

#### Load and clean the birthweight dataset. 

```{r}
baby_df = 
  read_csv("./data/birthweight.csv", na = c("", "NA", "Unknown")) %>% 
  mutate(
    babysex = as.factor(babysex),
    frace = as.factor(frace),
    frace = recode(frace, "1" = "White", "2" = "Black", "3" = "Asian", "4" = "Puerto Rican", "8" = "Other"),
    malform = as.factor(malform),
    mrace = as.factor(mrace),
    mrace = recode(mrace, "1" = "White", "2" = "Black", "3" = "Asian", "4" = "Puerto Rican", "8" = "Other"),
    smoken = as.factor(smoken),
    fincome = fincome * 100
  )
```

### Fit the models

We will examine three models to assess birthweight: 

1. Crude model examining length at birth and gestational age
2. Interactions model examining head circumference, length, sex, and all interactions between these predictors (including three-way interaction)
3. Full model examining head circumference, length, sex, mother's race, mother's pre-pregnancy BMI, family income, number of cigarettes smoked per day during pregnancy

**1. Crude model including main predictors**

- `blength`: baby's length at birth (cm)
- `gaweeks`: gestational age in weeks

```{r}
crude_model = lm(bwt ~ blength + gaweeks, data = baby_df)

crude_model %>% 
  broom::glance()
```

**2. Interaction model including variables**

- `bhead`: head circumference
- `blength`: baby's length at birth (cm)
- `babysex`: baby's sex

```{r}
inter_model = lm(bwt ~ bhead + blength + babysex + bhead * blength + bhead * babysex + blength * babysex, data = baby_df)
```

**3. Full model including variables**

- `bhead`: head circumference
- `blength`: baby's length at birth (cm)
- `babysex`: baby's sex
- `mrace`: mother’s race
- `ppbmi`: mother’s pre-pregnancy BMI
- `fincome`: family monthly income
- `smoken`: average number of cigarettes smoked per day during pregnancy

```{r}
full_model = lm(bwt ~ bhead + blength + babysex + mrace + ppbmi + fincome + smoken, data = baby_df)
```


### Plot the models

**1. Model of residuals vs. predicted values for crude model**

```{r}
crude_plot = 
  baby_df %>% 
  modelr::add_residuals(crude_model) %>% 
  modelr::add_predictions(crude_model) %>% 
  ggplot(aes(x = pred, y = resid)) +
  geom_point(aes(alpha = 0.5)) +
  labs(
    title = "Residuals vs. predicted values for crude model",
    x = "Predicted values",
    y = "Residuals"
  )

crude_plot
```

**2. Model of residuals vs. predicted values for interactions model**

```{r}
inter_plot = 
  baby_df %>% 
  modelr::add_residuals(inter_model) %>% 
  modelr::add_predictions(inter_model) %>% 
  ggplot(aes(x = resid, y = pred)) +
  geom_point(aes(alpha = 0.5)) + 
  labs(
    title = "Residuals vs. predicted values for interaction model",
    x = "Predicted values",
    y = "Residuals"
  )

inter_plot
```

**3. Model of residuals vs. predicated values for full model**

```{r}
full_plot = 
  baby_df %>% 
  modelr::add_residuals(full_model) %>% 
  modelr::add_predictions(full_model) %>% 
  ggplot(aes(x = resid, y = pred)) +
  geom_point(aes(alpha = 0.5)) + 
  labs(
    title = "Residuals vs. predicted values for full model",
    x = "Predicted values",
    y = "Residuals"
  )

full_plot
```

#3 Cross-validation of models

We will compare the full model with the crude and interaction models by:

- Creating test-training pairs
- Unpacking the pairs
- Assessing prediction accuracy
- Plotting the RMSE's 

```{r}
cv_df = 
  crossv_mc(baby_df, 100) %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  ) %>% 
  mutate(
    full_model = map(train, ~lm(bwt ~ bhead + blength + babysex + mrace + ppbmi + fincome + smoken, data = baby_df)),
    crude_model = map(train, ~lm(bwt ~ blength + gaweeks, data = baby_df)),
    inter_model = map(train, ~lm(bwt ~ bhead + blength + babysex + bhead * blength + bhead * babysex + blength * babysex, data = baby_df))
  ) %>% 
  mutate(
    rmse_full = map2_dbl(full_model, test, ~rmse(model = .x, data = .y)),
    rmse_crude = map2_dbl(crude_model, test, ~rmse(model = .x, data = .y)),
    rmse_inter = map2_dbl(inter_model, test, ~rmse(model = .x, data = .y))
  )
  
cv_plot = 
  cv_df %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model",
    values_to = "rmse", 
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + 
  geom_violin()

cv_plot
```

When comparing the violin plots of the root mean squared errors, it seems the full model which I hypothesized is the most predictive of the variable `birthweight` out of the three models since it has the lowest RMSE. 






